---
title: "msdrCourseraCapstone vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{msdrCourseraCapstone_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
#add this to run in development: 
#devtools::load_all(".")
```

# Overview 

This package was developed for the capstone project of the Mastering Software Development in R Coursera Specialization. The package provides functions for visualizing the NOAA Significant Earthquakes dataset. 

## Loading and cleaning the dataset 

```{r setup}
library(msdrCourseraCapstonePackage)
library(magrittr)
library(dplyr)
library(lubridate)
library(ggplot2)

data_path <- system.file("extdata", "signif.txt", package = "msdrCourseraCapstonePackage")

eq_data_raw <- read.delim(data_path)

eq_data <- 
  eq_data_raw %>% 
  eq_clean_data() %>% 
  eq_location_clean()
```

## Visualizing the data with timelines 

- First subset the data  

```{r visualization_1}
eq_data_plot <- eq_data %>% filter(COUNTRY %in% c("CANADA", "USA"))
```

- A simple example:  

```{r visualization_2, fig.width=7, fig.height=2}
ggplot(eq_data_plot, aes(x = DATE)) + 
  geom_timeline() + 
  theme_eq
```

- An example that uses all of the aes: 

```{r visualization_6, warning=FALSE, fig.width = 7, fig.height=8}
# use size and color of points to represent magnitude and number of casualties, and set alpha to 0.3
ggplot(eq_data_plot, aes(x = DATE, y = COUNTRY, min_date = as.Date("1900-01-01"), max_date = as.Date("2000-01-01"), size = EQ_PRIMARY, color = TOTAL_DEATHS)) + 
  geom_timeline(alpha = 0.3) + 
  geom_timeline_label(aes(n_max = 3, label = LOCATION_NAME)) + 
  theme_eq
```


## Visualizing the data with maps

- a map example 

```{r map_1, fig.width=7, fig.height=7} 
eq_data %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  eq_map(annot_col = "DATE")
```


- a map example with annotations created using eq_create_label(): 

```{r map_2, fig.width=7, fig.height=7}
eq_data %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```
